---
title: "Multi-level modeling"
---

```{r setup}
pacman::p_load(tidyverse, broom, lmerTest, sjPlot)
ess_work <- read_rds("data/ess_work.RDS")
```

## Your Turn 1

* Note that we are switching to a new dataset - ess_work, load it above
* Add a linear regression line to the plot below
* Then copy the plot and add color = cntry to the mapping - what do you observe about the relationship between health and happiness?

```{r}
ggplot(ess_work, aes(health, happy)) + 
  geom_jitter(alpha = .1) 

```

## Your Turn 2

* Test whether health predicts happiness, using lm
* Test this again, using lmer and adding a random slope per country (`cntry`) (syntax: `lmer(IV ~ DV + (1|group), data)`)
* Compare the results


```{r}

```



## Your Turn 3

* Take the lmer code from step 2 and add random slopes for `health` per country
* Add gender (`gndr`) as a fixed effect (just like a predictor in lm)
* Have a look at the summary() - is gender significant? What else has changed?
* Use ranova() to check whether the random slopes are significant.    

```{r}

```

## Visualise random effects

* Save your model from the last step into an object (`mod1 <- lmer(...`)
* Run the following code to see the random effects plot

```{r}
plot_model(mod1, type = "re") + 
  geom_hline(yintercept = 0)
```

## Your Turn 4: Bringing it all together + cross-level interactions


* Filter `ess_work` to create a dataset that only includes women (`gender` variable).
* Create an lmer model predicting working hours (`wkhtot`) depending on whether they have a `young_child` and include random intercepts per country.
* Add `gender_bias_men` as a predictor that interacts with `young_child` (remember: interactions are added by using * rather than +)
* Have a look at the summary: is there evidence for that interaction?
* Now add a random slope for `young_child` - what changes?


```{r}

```

NB: For the `gender_bias_men`-variable, I've calculated how much gender bias men in each country show based on the `aftjbyc` variable in the ESS. It is the difference in how much they approve of working fathers vs. working mothers of children below 3 years old.


***

# Take Aways

* Use multilevel models when observations are not independent

* `lme4` package (with `lmerTest` if you want p-values) can estimate such models

* Random intercepts for `group` are specified by adding `+ (1 | group)` to the formula

* Random slopes for `pred` in each `group` are specified by adding `+ (pred | group)` to the formula (note that the intercept is automatically added as well, no need to type the 1 if you have random slopes. It can be suppressed, but is usually wanted).

* Use `plot_model()` from the sjPlot package to plot any regression model. type = "re" gives random effects, and type = "int" gives interaction plots.