---
title: "Scales, mediation and moderation"
---

```{r setup}
pacman::p_load(tidyverse, broom, lavaan, processR, sjPlot, interactions)
pacman::p_load_gh("lukaswallrich/timesaveR")

ess_work <- read_rds("data/ess_work.RDS")

# We will also work with some data from my PhD - this code downloads it directly from GitHub and 
# prepares it for analysis. You can read through the code to practice your understanding, but don't need to do anything with it

contact <- read_csv("https://raw.githubusercontent.com/LukasWallrich/PhD_thesis/master/Chapter%203%20-%20mediation%20UK/analysis/AllResponses.csv", show_col_types = FALSE)

contact <- contact %>% 
  # Filter White respondents
  filter(Ethnicity %in% c(10, 11, 12)) 

# Create scales
scales <- list(
  affect_black = c("AffBlWarm", "AffBlNeg"),
  bystander_intervention = c(paste0("BSSlur", 2:7)),
  positive_contact = c("WhExpBlComplimented", "WhExpBlBefriended",
                "WhExpBlMadeFeelWelcome", "WhExpBlSupported", "WhExpBlHelped")
)

scale_results <- make_scales(contact, scales, reversed = TRUE, print_hist = FALSE, harmonize_ranges = FALSE)

contact <- bind_cols(contact, scale_results$scores)
```


## Your Turn 1

* Import the contact data (run the read_csv line above that gets it directly from my PhD repository).
* Run two `lm()` models to test the preconditions for testing whether the association between bystander_intervention and positive_contact (intergroup contact) is mediated by affect_black (affect towards Black people)
* Namely: is Y predicted by X and is Y predicted by M

```{r}


```

## Your Turn 2

* Type the lavaan model into the blank (___), using the correct variable names
* Run the whole block and interpret the output (NB: this will take ~20 seconds)


```{r}
model_code <- ("
 _____
               ")

set.seed(170524)

sem(model_code, contact, se = "bootstrap", bootstrap = 1000) %>% 
  summary(ci = TRUE)

```


## Your Turn 3

* Test whether education (`eduyrs`) moderates the relationship between health and happiness (in the `ess_work` data)
  - Run the `lm()` model testing whether health interacts with income in predicting happiness
  - Put the result into the `plot_model()` function with `type = "int"` and `mdrt.values = "meansd"` - if it does not show what you want, try to switch the order of the predictors in `lm()`.

* There are some extreme outliers in `eduyrs` - to avoid issues, we will simply windsorize them (set them to the 1st and 99th percentile), in a real analysis, this would require further justification and exploration. Here, just run that code before the `lm()` model.

NB: `plot_model()` sometimes struggles with labelled data - therefore, it's worth using zap_labels() here. Also, note that thist just returns a ggplot-object, so that you can format it easily.

```{r}
ess_work <- ess_work %>% mutate(eduyrs = eduyrs %>%
                                  pmax(quantile(., .01, na.rm = TRUE), .) %>% 
                                  pmin(quantile(., .99, , na.rm = TRUE), .))


mod_mod <- ess_work %>% haven::zap_labels() %>% lm(____, data = .)

plot_model(______) + 
  labs(title = "Predicted happiness by health and education") +
  jtools::theme_apa()
```


## Reporting simple slopes and Johnson-Neyman intervals

You can use the code below to get all details needed to report simple slopes. `?interactions::sim_slopes` reveals many customisation options.

```{r}
interactions::sim_slopes(mod_mod, pred = "health", modx = "eduyrs")
```

## Your turn 4

* Fill in the blanks and try to understand what's going on
* Run the code and compare the output to the simple slopes above

```{r}

model_code <- ("
  #Regression model
  happy ~ a * health + b * eduyrs + c * health: eduyrs
  
  #Ingredients for simple slopes
  eduyrs ~ ______ * 1
  eduyrs ~~ ______ * eduyrs
  
  #Simple slopes
  SDbelow := a + c * (____ - ____(______))
  Mean := a + c * ______
  SDabove:= a + c * _____
")


sem(model = _____, data = ____) %>% 
  summary()

```


***

# Take Aways

* Mediation and moderation models can be coded in `lavaan` - highly flexible, and learning then also transfers to  SEM and other path modeling
* Coding is not trivial - but no need to start from scratch. Many code examples can be found - most comprehensively in the processR package/[web app](https://cardiomoon.shinyapps.io/processR/)
* Simpler tools are available - `JSmediation` package looks promising, and often for moderation, `lm()` and the simple decomposition functions shown above are sufficient.