---
title: "Visualize Data"
---

## Your Turn 0

Add a setup chunk that loads the tidyverse packages and the dataset. For the dataset, use 
ess_mini <- read_rds("data/ess_mini.RDS")
*after* you have loaded the tidyverse (because read_rds comes from the tidyverse). 
Remember that the code needs to be inside a code block to work.



## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!
```{r}

```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph, inside the aes() function. Experiment.  

```{r}
#To start, copy the code from Turn 1
```

## Have a look

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, discuss with a partner)

```{r}
# Makes a plot that the commands below will modify
p <- ggplot(data = ess_mini) +
  geom_point(mapping = aes(x = height, y = weight), alpha = 0.1)

p + facet_grid(rows = vars(country))
p + facet_grid(cols = vars(country))
p + facet_grid(rows = vars(country), cols = vars(gender))
p + facet_wrap(vars(main_activity))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet and/or try your best guess.

```{r}
ggplot(data = ess_mini) +
  geom_point(mapping = aes(country, working_hours, color = gender))
```

## Your Turn 4

Make a histogram of the `working_hours` variable from `ess_mini`. Hint: do not supply a y variable.

Bonus: facet it by country

```{r}

```

## Your Turn 5

Use the help page for `geom_histogram` to make the bins 5 units wide. To get there, just type `?geom_histogram` in the Console.

Bonus: use the help page for `facet_wrap` to allow for different scales for the y-axis in each subplot ('free' the scales).

```{r}

```

## Your Turn 6

Make a bar chart of `main_activity` colored by `country`. Note that `geom_bar` accepts both a `color` and a `fill` aesthetic. Try them out to see the difference.

```{r}

```

## Quiz

What will this code do?

```{r}
ggplot(ess_mini) + 
  geom_point(aes(height, weight)) +
  geom_smooth(aes(height, weight))
```

## Making plots interative

ggplot2 can combine with plotly to make graphs interactive - which could be great for graphs with relatively few points and lots of information. If you want to explore this further, have a look at the [plotly webpage](plot.ly/ggplot2/)

```{r}
p <- ggplot(ess_mini, aes(height, weight, shape = country, color = gender)) + 
  geom_point() +
  geom_smooth()

pacman::p_load(plotly)
ggplotly(p)
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

NB:You can also change data in each geom_* function, or provide the mappings overall in the ggplot() function - but the template above is a good starting point.

Hint: if you want your plots to look more "journal-ready" you can add `+ jtools::theme_apa(legend.use.title = TRUE)` to use that theme provided by the `jtools` package. For that, you need to install that package, e.g. by running `install.packages("jtools")` in the Console. Alternatively, there are many other themes available in the `ggplot2` ecosystem, some directly within ggplot (e.g., `theme_minimal()`), others in packages (e.g., `ggthemes::theme_economist()`).

## Complete example - the gapminder graph

This is an example pulling together many different functions - you certainly don't need to understand all of that, but might find it helpful as a starting point for further exploration.

```{r}
pacman::p_load(dslabs,  #Teaching package that contains the gapminder dataset
               jtools)  #For the theme_apa function
               
#Filter dataset to contain 4 years - we'll learn how to do that next
gapminder <- gapminder %>% filter(year %in% c(1960, 1980, 2000, 2010)) %>% 
  mutate(gdp_per_capita = gdp/population)

ggplot(gapminder, aes(x=gdp_per_capita, y=life_expectancy, col=continent, size=population/10^6)) +
                                                                          #^ Population converted to millions
  geom_point(alpha = .5)  + 
  scale_x_log10() + 
  scale_size(range=c(.1, 15)) +  #Increased range makes largest points larger
  scale_color_brewer(type = "qual") + 
  facet_wrap(vars(year)) + 
  jtools::theme_apa(legend.use.title = TRUE) + 
  labs(title = "The past 50 years: Greater wealth and greater health",
       x = "GDP per capita (USD/year)",
       y = "Life expectancy (years)",
       color = "Continent",
       size = "Population (M)")
```

