---
title: "Visualize Data"
---

## Your Turn 0

Add a setup chunk that loads the tidyverse packages and the dataset. For the dataset, use `ess_mini <- read_rds(“ess_mini.RDS)` *after* you have loaded the tidyverse.

```{r setup}
pacman::p_load(tidyverse)
ess_mini <- read_rds("data/ess_mini.RDS")
```


```{r}
ess_mini
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = ess_mini) +
  geom_point(mapping = aes(x = height, y = weight))
```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = ess_mini) +
  geom_point(mapping = aes(x = height, y = weight, color = gender, shape = country))
```

## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
# Makes a plot that the commands below will modify
p <- ggplot(data = ess_mini) +
  geom_point(mapping = aes(x = height, y = weight), alpha = 0.1)

p + facet_grid(rows = vars(country))
p + facet_grid(cols = vars(country))
p + facet_grid(rows = vars(country), cols = vars(gender))
p + facet_wrap(vars(main_activity))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = ess_mini) +
  geom_point(mapping = aes(country, working_hours, color = gender))
```

## Your Turn 5

Make a histogram of the `working_hours` variable from `ess_mini`. Hint: do not supply a y variable.

```{r}
ggplot(ess_mini) + 
  geom_histogram(aes(x = working_hours)) + 
  facet_wrap(vars(country))

ggplot(ess_mini) + 
  geom_histogram(aes(x = working_hours, fill = working_hours %% 5 == 0), binwidth = 1) +
  labs(fill = "Divisible by 5")

ggplot(ess_mini) + 
  geom_histogram(aes(x = working_hours), binwidth = 5) 


ggplot(ess_mini) + 
  geom_histogram(aes(x = working_hours)) + 
  facet_wrap(vars(country))

```

## Your Turn 5

Use the help page for `geom_histogram` to make the bins 5 units wide.
Bonus: use the help page for `facet_wrap` to free the scales for the y-axis.

```{r}
ggplot(ess_mini) + 
  geom_histogram(aes(x = working_hours), binwidth = 5) + 
  facet_wrap(vars(country), scales = "free_y")


ggplot(ess_mini) + 
  stat_ecdf(aes(x = working_hours, color = country, linetype = gender))

```

## Your Turn 6

Make a bar chart of `main_activity` colored by `country`. Note that `geom_bar` accepts both a `color` and a `fill` aesthetic - experiment to see the difference and decide which you need. 

```{r}
ggplot(ess_mini) +
  geom_bar(aes(y = main_activity, fill = gender))

ggplot(ess_mini) +
  geom_bar(aes(y = main_activity, fill = gender), position = "fill")

```

## Quiz

What will this code do?

```{r}
ggplot(ess_mini) + 
  geom_point(aes(height, weight)) +
  geom_smooth(aes(height, weight))

ggplot(ess_mini, aes(height, weight)) + 
  geom_point(aes(color = gender)) +
  geom_smooth()

ggsave("g.png")

```

## Quiz

What is different about this plot? Run the code!

```{r}
p <- ggplot(ess_mini, aes(height, weight)) + 
  geom_point() +
  geom_smooth()

library(plotly)
ggplotly(p)
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

Hint: if you want your plots to look more "journal-ready" you can add `+ jtools::theme_apa(legend.use.title = TRUE)` to use that theme provided by the `jtools` package. For that, you need to install that package, e.g. by running `install.packages("jtools")` in the Console.

## Complete example - the gapminder graph

This is an example pulling together many different functions - you certainly don't need to understand all of that, but might find it helpful as a starting point for further exploration.

```{r}
pacman::p_load(dslabs, jtools, scales) #Teaching package that contains the gapminder dataset

#Filter dataset to contain 4 years - we'll learn how to do that next
gapminder <- gapminder %>% filter(year %in% c(1960, 1980, 2000, 2010)) %>% 
  mutate(gdp_per_capita = gdp/population)

ggplot(gapminder, aes(x=gdp_per_capita, y=life_expectancy, col=continent, size=population/10^6)) +
  #^ Population converted to millions
  geom_point(alpha = .5)  + 
  scale_x_continuous(trans = scales::log2_trans()) +
  #Log transformation moved here - that gives nicer labels
  scale_size(name = "Population (M)", range=c(.1, 15)) + 
  #Increased range makes largest points larger
  scale_color_discrete(name="Continent") + 
  facet_wrap(vars(year)) + 
  jtools::theme_apa(legend.use.title = TRUE) + 
  labs(title = "The past 50 years: Greater wealth and greater health",
       x = "GDP per capita (USD/year)",
       y = "Life expectancy (years)")
```


