---
title: "Models"
---

```{r setup}
pacman::p_load(tidyverse, broom)
pacman::p_load_gh("LukasWallrich/timesaveR")
```

## Your Turn 1

* Import "ess_mini.RDS" from the "data" folder and include that code in the setup chunk 
* Filter respondents from the UK (where country is "GB") and save in (`<-`) ess_mini_UK
* Create boxplot to show working hours for males and females
* Calculate mean working hours for males and females

```{r}

```


## Your Turn 2

* Run a t.test to compare working hours to test whether working hours significantly differ between men and women in the UK.
* Read through the output - what do you see?

```{r}

```

## Your Turn 3

Test whether mean happiness differs between (pairs of) countries.

```{r}

```

## Reading practice

How does this code work?

```{r}
ess_mini %>% 
    mutate(gender_country = paste(gender, country)) %>%
    pairwise_t_tests(height, gender_country, p.adjust.method = "none") %>% 
    filter(str_detect(var_1, "Male") & str_detect(var_2, "Male") |
              str_detect(var_1, "Female") & str_detect(var_2, "Female")) %>%
    arrange(desc(abs(cohens_d))) %>%
    select(var_1, var_2, apa)
```


## Your Turn 4

Test whether happiness and health are correlated

Calculate both Pearson’s r and Spearman’s rho.
Remember ? to get details on arguments: ?cor.test

```{r}

```

## Your Turn 5

Explore the relationship between body mass index and self-reported health.

* Add a bmi variable to the ess_mini dataset (which dplyr verb was that yet again?). bmi is defined as the weight divided by the height in m squared (e.g., it might be 80/(180/100)^2 for someone who weighs 80 kg and is 180 cm tall)

* Add `+ geom_smooth(method = "lm")` to ggplot code below and have a look at the result. What relationships do you see?

```{r}

ess_mini <- ____

ess_mini %>%
     ggplot(aes(____)) +
     geom_jitter(alpha = ____) + 
     _____

lm(____) %>% summary(____)

```


## Your Turn 6

Check whether our regression model meets the key assumptions

* Save your health ~ bmi + gender lm-model into `mod`
* Install and load the performance and see packages (`pacman::p_load` takes care of installing *and* loading)
* Test the model with `check_model(mod)`
* Discuss the results with a partner

```{r}


```

***

# Take Aways

* Use `t.test()`, `cor.test()` and `lm()` to conduct simple statistical tests

* Specify models with formulae where `~` means 'is predicted by'

* Consider `pairwise_t_tests()`, `cor_matrix()` and `lm_std()` from `timesaveR` to facilitate common 'next steps'

* Use R to create publication ready tables - try out `report_cor_table()`, `report_lm_with_std()` and `report_cat_vars()` from `timesaveR` or more general functions from the `modelsummary` package

* Use `glance()`, `tidy()`, and `augment()` from the **broom** package to return model values in a data frame.

* Use `check_model()` from the `performance `package to quickly test how an `lm()`-model performs with respect to they assumptions made for linear regression
