---
title: "Importing data and resolving issues"
---

```{r setup}
pacman::p_load(tidyverse)
pacman::p_load(sjPlot)
pacman::p_load_gh("lukaswallrich/timesaveR")

```

## Your Turn 1

* Add code to the setup chunk to import "qualtrics_example.sav" from the "data" folder. Use `haven::read_sav()`) and assign it to `survey_data`
* Have a look at the data by clicking on it in the environment
* Run `sjPlot::view_df(survey_data)` and briefly explore the output
* Run `count(survey_data, Q2.2)` and look at the output
* Run the ggplot code below and look at the output and warning message

```{r}

ggplot(survey_data, aes(Q2.2)) + geom_bar()
```

## Your Turn 2

* Make a new factor variable gender based on Q2.2 (`as_factor()`)
* Rename Q2.1 to age
* Replace Q1.2_ in variable names with "consent". Note the order of arguments for str_replace: (string, pattern_to_look_for, replacement), as in the example below.

```{r}
_____
_____

names(survey_data) <- str_replace(names(survey_data), "_DO_", "_order_")

```


## Your Turn 3

* Complete the code to calculate the three scales:
  - UCLA loneliness: Q59_1 to Q59_20
  - Warwick-Edinburgh wellbeing: Q60_1 to Q60_11
  - Covid_prevention: Q61_1 to Q61_12
  
* Calculate the scale scores. Use ?make_scales to check how to set `reversed` (hint: TRUE or FALSE), then run that line.

* Have a look at the histograms and at `scales$descriptives`

* Bind the scores (`scales$scores`) to the survey_data

NB: Much of the code is there - just fill in the ___

```{r}
scales_items <- list(
  loneliness = paste0("Q59_", 1:20),
  wellbeing = _____,
  ____
  )

scales <- make_scales(ess_health, scales_items, reversed = ___)

survey_data <- survey_data %>% bind_cols(_____)

```

## Your Turn 4

Each line below has a few mistakes - most mean that the code won't run, others result in meaningless results. See how many you can fix within 5 minutes. Remember that you can try running a piece of code by highlighting it and pressing Ctrl + Enter.

```{r}
#This is correct - we need some data
ess_mini <- read_rds("data/ess_mini.RDS")

###
# Transforming data
###

# I want to know the number of respondents and median working hours per country
# 3 mistakes
ess_mini %>% 
  by(country) %>% 
  summarise(median_work <- median(working_hours, na.rm = TRUE), N = count())

# I want to add up two variables - assume that height + weight is meaningful
# 1 mistake that does not lead to an error - but to unexpected and meaningless results
# So make sure to look at the output
ess_mini %>% mutate(new_var = sum(height, weight, na.rm = TRUE))

###
# Plotting data
###

# I want to show the link between happy and working hours among women, 
# with a scatter plot overlaid by linear regression lines coloured by country
# 2-3 mistakes per line
ess_mini %>% filter(gender = Female) %>%
  ggplot(aes(happy, working_hours) %>%
    geom_point() + geom_smooth(color = "country")       

###
# Modeling
###    

# I want to test whether loneliness predicts lower wellbeing
# For that, I need a p-value
# Two changes needed to get there.
survey_data %>%  
  lm(wellbeing ~ loneliness)

```

***

# Take Aways

* Use `haven::read_sav()` to import SPSS files (and similar functions for SAS and Stata)

* Note that R is not great with such labeled data - make sure you have the codebook elsewhere (e.g., from `sjPlot::view_df()`) and convert variables as necessary, with `as_factor()` for categorical variables. `haven::zap_labels()` removes all labels from a dataset and can be useful when you run into trouble

* `timesaveR::make_scales()` can be helpful for making scales, including reverse coding and assessing internal consistency.

* Debugging code that is not doing what you expect it to do is the best way to learn :). ChatGPT and co can be very helpful there.
